---
title: "BATSSSSSSS"
author: "Tad Dallas"
output:
  pdf_document:
    toc: yes
  html_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 10
    highlight: tango
    theme: journal
    toc: yes
---





```{r eval=TRUE, echo=FALSE, comment="", message=FALSE}

## Weird path really only useful on Colin's machine
# dat <- read.csv('~/GitHub/cleanbats_betacov/clean data/BatCoV-assoc_compatible.csv', stringsAsFactors=FALSE)

# host <- read.csv('~/GitHub/cleanbats_betacov/clean data/Han-BatTraits_compatible.csv',stringsAsFactors=FALSE)


dat <- read.csv('virionette.csv', stringsAsFactors=FALSE)
host <- read.csv('Han-BatTraits.csv',stringsAsFactors=FALSE)
cites <- read.csv('Citations.csv',stringsAsFactors=FALSE)

set.seed(666)

```




# Load packages

```{r}

library(Amelia)
library(fields)
library(ROCR)
library(corpcor)
library(np)
library(plyr)
library(dplyr)

```



# Make some colors 
```{r}
ose1 <- rgb(85,108,17, m=255)
ose2 <- rgb(160,108,17, m=255)
ose3 <- rgb(114,132,56, m=255)
ose4 <- rgb(137,152,87, m=255)
```





```{r, eval=TRUE, echo=FALSE}

#' Plug and Play algorithm
#' Applies the plug and play algorithm to data on environmental covariates at presence and background using multivariate Gaussian distribution for f and f1
#'
#'  @param p covariate data at presence/occurrence points
#'  @param bgrd covariate data at background points
#'  @param method either 'classical' (default), 'robust', or 'regularized'
#'  
#'  @returns A list contain the means and covariances
#'  

pp.gauss <- function(p, bgrd, method='regularized', vbose = FALSE){
  mean.p <- colMeans(p)         # estimated mean of presence points
  mean.bgrd <- colMeans(bgrd)   # estimated mean of background points
  sigma.p <- stats::cov(p)             # estimated covariance of presence points
  sigma.bgrd <- stats::cov(bgrd)       # estimated covariance of background points

  if(method=='robust'){
    require(robust)
    p.est <- covRob(p)
    mean.p <- p.est$center        # robust estimated mean
    sigma.p <- p.est$cov          # robust estimated covariance of presence points
    bgrd.est <- covRob(bgrd)
    mean.bgrd <- bgrd.est$center  # robust estimated mean
    sigma.bgrd <- bgrd.est$cov    # robust estimated covariance of background points
  }
  if(method=='regularized'){
    require(corpcor)
    # robust estimated covariance of presence points
    p.est <- as.numeric(cov.shrink(p, verbose= vbose))  
    sigma.p <- matrix(as.numeric(p.est), nrow=sqrt(length(p.est)))  
    # robust estimated covariance of background points
    bgrd.est <- as.numeric(cov.shrink(bgrd, verbose = vbose))    
    sigma.bgrd <- matrix(as.numeric(bgrd.est), nrow=sqrt(length(bgrd.est))) # reformat
  }
  return(model=list(mean.p=mean.p, sigma.p=sigma.p, mean.bgrd=mean.bgrd, sigma.bgrd=sigma.bgrd))
}



#'  Takes a learned gaussian-plug-and-play and applies to a new point
#'  
#'  @param model a list object from pp.gauss
#'  @param x  a vector of points to classify
#'
#'  @returns A vector of relative suitabilities

predict.pp.gauss <- function(model, x){
  require(mvtnorm)
  f1 <- dmvnorm(x, mean=model$mean.p, sigma=model$sigma.p, log=TRUE)
  f0 <- dmvnorm(x, mean=model$mean.bgrd, sigma=model$sigma.bgrd, log=TRUE)
  return(S=exp(f1-f0))
}




#' Applies the plug and play algorithm to data on environmental covariates at presence and background using multivariate Gaussian distribution for f and f1
#'
#'  @param p covariate data at presence/occurrence points
#'  @param bgrd covariate data at background points
#'
#'  @returns model a list object containing f1 and f0 

pp.kde <- function(p, bgrd){
  require(np)
  f1 <- npudens(p, bwmethod='normal-reference')
  f0 <- npudens(bgrd, bwmethod='normal-reference')
  return(model=list(f1=f1,f0=f0))
}




#'  Takes a learned kde-plug-and-play and applies to a new point
#'  
#'  @param model a list object from pp.gauss
#'  @param x  a vector of points to classify
#'
#'  @returns A vector of relative suitabilities

predict.pp.kde <- function(model, x){
  require(np)
  f1 <- fitted(npudens(bws=model$f1$bws, edat=data.frame(x), bwmethod='normal-reference'))
  f0 <- fitted(npudens(bws=model$f0$bws, edat=data.frame(x), bwmethod='normal-reference'))
  return(S=f1/f0)
}

```


# The code below is used to calculate the number of estimated missing links from host-parasite networks


```{r eval=TRUE, echo=F}

#' Function to compute frequnecy of frequencies table from a list of records
#' Input: x (a list of unique records)
#' 
#'
#' @returns frequency of frequencies table 

FoF <- function(x){
  x.0 <- table(table(x))
  out <- cbind(as.numeric(names(x.0)),as.vector(x.0))
  out
}



#' Function to estimate and summarize missing classes
#' @param fof (a frequency-of-frequencies table)
#' 
#' @returns list object containing observed, estimated, and missing links

summarize <- function(fof){
  observed <- sum(fof[,2])
  ace <- ChaoLee1992(fof, t = 10, method = "all", conf = 0.95)
  number.missing <- ace$Nhat - observed
  fraction.observed <- observed/ace$Nhat[2:1]
  out <- list(observed=observed, estimated=ace$Nhat[2:1], number.mising=number.missing, fraction.observed=fraction.observed)
  out
}

```






---



# Host-virus associations


```{r eval=TRUE, echo=TRUE}

#' imputes missing data.
#'
#' @param listobject contains interaction matrix (intMat), host traits, parasite traits, and a points object that gets handed to pp.gauss
#' 
#' @returns the same listobject, now with missing host and parasite trait data imputed and pointsobject updated.

getImpute <- function(listobject, m=10, rem=FALSE, useMeans=TRUE){
	require(Amelia)
	require(vegan)

  htriz <- listobject$Hosts
	ptriz <- listobject$Parasites

  #remove columns with less than 75% data coverage
  no  <- which(colSums(is.na(htriz))/nrow(htriz) > 0.75)
	nop <- which(colSums(is.na(ptriz))/nrow(ptriz) > 0.75)
  if(length(no) > 0){
    htriz <- htriz[,-no]
  }
  if(length(nop) > 0){
    ptriz <- ptriz[,-nop]
  }

# remove columns with correlation coeff
  no2 <- which(cor(htriz, use='complete.obs')[,1] > 0.6)
  if(length(no2) > 1){
    no2 <- no2[-1]
		htriz = htriz[,-no2]
  }

  #remove rows without any data
	#hosts
  inds <- which(is.na(colSums(htriz)))
  if(rem==TRUE){
    cut=rep(0,nrow(htriz))
    for(i in seq_len(nrow(htriz))){
      if(all(is.na(htriz[i,]))){cut[i]=1}
    }
    htriz <- htriz[-which(cut==1), ]
  }
	#parasites
  indsP <- which(is.na(colSums(ptriz)))
  if(rem==TRUE){
    cut=rep(0,nrow(ptriz))
    for(i in seq_len(nrow(ptriz))){
      if(all(is.na(ptriz[i,]))){cut[i]=1}
    }
    ptriz <- ptriz[-which(cut==1), ]
  }

 #impute host traits
 if(any(is.na(colSums(htriz)))){
    helper <- which(colSums(htriz,na.rm=TRUE)>1)[1]
		tmp <- htriz[, unique(c(helper,inds))]
		if(useMeans==TRUE){
			means <- which(apply(tmp, 1, function(x){sum(is.na(x))})/ncol(tmp) > 0)
			tmp[means, ] <- colMeans(tmp, na.rm=TRUE)
			htriz2 <- tmp
		}else{
	    htriz2 <- amelia(x = tmp, m=m, p2s=0)$imputations[[m]]
		}
	  notInds <- seq(1,ncol(htriz),by=1)[-c(helper,inds)]
	  htriz3 <- cbind(htriz2, htriz[,notInds])
  }else{
    htriz2 <- amelia(x = htriz[, c(inds)], m=m, p2s=0)$imputations[[m]]
    notInds <- seq(1,ncol(htriz),by=1)[-c(inds)]
    htriz3 <- cbind(htriz2, htriz[,notInds])
  }


  #remake the points object
  intMat <- listobject$intMat
  pos <- which(intMat == 1, arr.ind=TRUE)
  neg <- which(intMat == 0, arr.ind=TRUE)
  ret <- rbind(pos,neg) 
  ret <- cbind(ret, c(rep(1, nrow(pos)), rep(0,nrow(neg)))) 
  colnames(ret) <- c('parasite', 'host', 'presence')

  pt <- ptriz[ret[,2],]
  ht <- htriz3[ret[,1],]
	ht <- ht[,-1]

 	ptobj <- data.frame(ret, ht, pt, stringsAsFactors=TRUE)
	return(list(intMat=intMat, Hosts=htriz3, 
		Parasites=ptriz, pointsObject=ptobj))
}


```



```{r eval=TRUE, echo=TRUE}

#' Train pp.gauss model on data
#'
#' @param pointsobject input to pp.gauss function
#' 
#' @returns list object of host and parasite information, model predictions, and AUC value

getReal <- function(pointsobject, cv=TRUE, folds=5){
  require(ROCR)
  # all the data divided into presence (p) and background (bg)
  p <- pointsobject[(pointsobject[,'presence'] == 1), 4:ncol(pointsobject)]
  bg <- pointsobject[,4:ncol(pointsobject)]
  if(cv==TRUE){
		# Take a random sample and exclude it
		ints <- sample(1:nrow(pointsobject), nrow(pointsobject)/folds)
		red <- pointsobject[-ints,]
		p.red <- red[(red[,'presence'] == 1), 4:ncol(pointsobject)]
		bg.red <- red[, 4:ncol(pointsobject)]
		# Train the model on a portion of the data
		cvmodel <- pp.gauss(p=p.red, bgrd=bg.red, vbose=FALSE)
		#... and predict on the holdout set
		red.preds <- predict.pp.gauss(cvmodel, x=bg[ints,])
		ret <- data.frame(presence=pointsobject[ints,'presence'], suitability=red.preds)
		hosts <- pointsobject[ints,'parasite']; parasites=pointsobject[ints,'host']
  }

  if(cv==FALSE){
    fullmodel <- pp.gauss(p=p, bgrd=bg, vbose=FALSE)
    redpreds <- predict.pp.gauss(fullmodel, x=bg)
    ret <- data.frame(presence=pointsobject[,'presence'], suitability=redpreds)
    hosts <- pointsobject$parasite; parasites=pointsobject$host
  }
  #auc calculation
  pred <- prediction(predictions=ret[,2], labels=ret[,1])
  acc <- as.numeric(performance(pred, 'auc')@y.values)
  return(list(hostInds=hosts, parasiteInds=parasites, predictions=ret, auc=acc))
}

```






```{r eval=TRUE, echo=F, message=FALSE, warning=FALSE}

makePoints <- function(intmat, HostTraits, ParasiteTraits){
  ret <- matrix(0,ncol=3, nrow=(nrow(intmat)*ncol(intmat))) 
  colnames(ret) <- c('row', 'column', 'presence')
  negInts <- which(intmat == 0, arr.ind=TRUE)
  posInts <- which(intmat > 0, arr.ind=TRUE)
  ret[,1:2] <- rbind(negInts, posInts)
  ret[,3] <- c(rep(0, nrow(negInts)), rep(1, nrow(posInts)))
  ret2 <- cbind(ret,HostTraits[(ret[,1]),], ParasiteTraits[(ret[,2]),])
  return(list(intMat=intmat, Hosts=HostTraits, Parasites=ParasiteTraits, pointsObject=ret2))
}


```







```{r eval=TRUE, echo=F, message=FALSE, warning=FALSE}
expandCol <- function(dat, colName=NULL){
	namez <- na.omit(unique(dat))
	ret <- sapply(namez, function(x){as.integer(grepl(x, dat))}) 
	colnames(ret) <- paste(colName, '.', namez, sep='')
	return(as.data.frame(ret))
}

```











```{r eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}


datBats <- filter(dat, host_order=='Chiroptera')

## interaction matrix
intMatFull <- as.data.frame.matrix(table(datBats$host_species, datBats$virus_genus))


### created features
numParasites <- rowSums(intMatFull > 0)
numHosts <- colSums(intMatFull > 0)

h <- data.frame(host_species=rownames(intMatFull))
host$host_species <- paste(host$MSW05_Genus, host$MSW05_Species, sep=' ')
h <- left_join(h, host, by='host_species')
h <- left_join(h, cites, by=c('host_species'='name'))
h$X <- NULL
h$Pan <- NULL






### forming data.frame objects
host2 <- data.frame(
	expandCol(h$MSW05_Genus, 'hGenus'), 
	expandCol(h$ForStrat.Value, 'hForStratValue'),
	expandCol(h$IUCN , 'hIUCN '),  
	h[,c(6:60, 62:68)], 
	numParasites
)

host2$X12.2_Terrestriality <- NULL
host2$Diet.Vunk<- NULL
host2$Diet.Scav <- NULL
host2$Activity.Nocturnal<- NULL 
host2$X12.1_HabitatBreadth<- NULL
host2$Diet.Vfish<- NULL
host2$Diet.Nect <- NULL
host2$Diet.Seed<- NULL
host2$Diet.PlantO<- NULL
host2$urbc_1970AD<- NULL
host2$urbc_1980AD<- NULL
host2$urbc_1990AD<- NULL
host2$urbc_2000AD<- NULL
host2$urbc_2005AD <- NULL
host2$hFamily.DELPHINIDAE <- NULL
host2$ForStrat.Value <- NULL
host2$IUCN <- NULL


# drop these too.
host2$BodyMass.value <- NULL
host2$PET_mean_mm <- NULL
host2$HuPopDen_5p_n.km2 <- NULL
host2$HuPop_Den_min_n.km2 <- NULL




virus2 <- datBats %>%
	group_by(virus_genus) %>%
	summarise(nHostSp=length(unique(host_species)),
		nHostFam=length(unique(host_family)),
		nHostOrder=length(unique(host_order)))

virus2$virus_genus <- NULL

## Points object
hvDat <- makePoints(intMatFull, host2, virus2)

```



```{r}
writeLinks <- function(dat, model, fileName='DallasPredictions.csv'){
	linksDF <- data.frame(host=rownames(dat[[1]])[model[[1]]],
		virus = colnames(dat[[1]])[model[[2]]], 
		presence = model[[3]]$presence,
		suitability = model[[3]]$suitability)
	linksDF <- linksDF[order(linksDF$suitability, decreasing=TRUE),]
	ret <- linksDF[which(linksDF$virus == 'Betacoronavirus'), ]
	write.csv(ret, file=fileName)
}
```










```{r eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

hvDati <- getImpute(hvDat, m=5, useMeans=TRUE)

```








3.       Filter (host_order = CHIROPTERA) > PNP with trait approach (no latent) >  include citation counts > Return a csv of bat species and probability of interaction with virus_genus = ‘Betacoronavirus’ as outcome variable > DallasBatsCitations.csv


```{r eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

hvOut1 <- getReal(hvDati[[4]], cv=FALSE)
writeLinks(dat=hvDati, model=hvOut1, fileName='DallasBatsCitations.csv')

```





4.       Filter (host_order = CHIROPTERA) > PNP with trait approach (no latent) >  don’t include citation counts > Return a csv of bat species and probability of interaction with virus_genus = ‘Betacoronavirus’ as outcome variable > DallasBatsUncorrected.csv


```{r eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
hvDati[[2]]$cites <- NULL
hvOut2 <- getReal(hvDati[[4]], cv=FALSE)
writeLinks(dat=hvDati, model=hvOut1, fileName='DallasBatsUncorrected.csv')
```












































### The approach we took in the Open Science article (just use dissimilarity values)


```{r}
#' Format the data from a list of interaction matrices to a data.frame 
#' that can be input to the plug and play model.
#'
#'
#' @param intMatList list of interaction matrices
#' @param year year of prediction
#' @param lagYears number of years which to train the model on
#' @param distFun how to calculate dissimilarity values
formatPPobject <- function(intMat,
	distFun='bray'){
	require(plyr)
	require(dplyr)
	locSim <- rowMeans(as.matrix(vegan::vegdist(intMat)))	
	pathSim <- rowMeans(as.matrix(vegan::vegdist(t(intMat))))
	simMat <- locSim %*% t(pathSim)	
	locDF <- data.frame(host = rownames(intMat), locSim)
	pathDF <- data.frame(pathogen = colnames(intMat), pathSim)
	df <- expand.grid(rownames(intMat), colnames(intMat))
	colnames(df) <- c('host', 'pathogen')
	df$presence <- as.vector(intMat > 0)+0 
	df$similarity <- as.vector(simMat)
	df <- left_join(df, pathDF, by='pathogen')
	df <- left_join(df, locDF, by='host')
	return(df)
}

```





```{r eval=TRUE, echo=TRUE}
#' Train pp.gauss model on data
#'
#' @param pointsobject input to pp.gauss function
#' @param cv (boolean) cross-validate the model?
#' @param folds number of cross-validation folds
#' @param newData data to predict on (default=NULL)
#' 
#' @returns list object of host and parasite information, model predictions, and AUC value
getReal2 <- function(pointsobject, cv=FALSE, folds=5, newData=NULL){
  require(ROCR)
  # all the data divided into presence (p) and background (bg)
  p <- pointsobject[(pointsobject[,'presence'] >= 1), 4:ncol(pointsobject)]
  bg <- pointsobject[,3:ncol(pointsobject)]
  if(cv==TRUE){
		# Take a random sample and exclude it
		ints <- sample(1:nrow(pointsobject), nrow(pointsobject)/folds)
		red <- pointsobject[-ints,]
		p.red <- red[(red[,'presence'] >= 1), 4:ncol(pointsobject)]
		bg.red <- red[, 3:ncol(pointsobject)]
		# Train the model on a portion of the data
		cvmodel <- pp.gauss(p=p.red, bgrd=bg.red, vbose=FALSE)
		#... and predict on the holdout set
		red.preds <- predict.pp.gauss(cvmodel, x=bg[ints,])
		ret <- data.frame(presence=pointsobject[ints,'presence'], suitability=red.preds)
		hosts <- pointsobject[ints,'host']
		parasites <- pointsobject[ints,'pathogen']
  }
  if(cv==FALSE){
    fullmodel <- pp.gauss(p=p, bgrd=bg, vbose=FALSE)
		if(!is.null(newData)){
			redpreds <- predict.pp.gauss(fullmodel, x=newData[,4:ncol(newData)])
		  ret <- data.frame(presence=newData[,'presence'], suitability=redpreds)
		  hosts <- newData$host
			parasites <- newData$pathogen
		}else{
			redpreds <- predict.pp.gauss(fullmodel, x=bg)
		  ret <- data.frame(presence=pointsobject[,'presence'], suitability=redpreds)
		  hosts <- pointsobject$host
			parasites <- pointsobject$pathogen
		}
  }
	tss <- function(x){
		max(unlist(performance(x, 'sens')@y.values)+ 
			unlist(performance(x, 'spec')@y.values) - 1)
	}
  #auc calculation
  pred <- prediction(predictions=ret[,2], labels=ret[,1])
  acc <- as.numeric(performance(pred, 'auc')@y.values)
	return(list(hostInds=hosts, parasiteInds=parasites, predictions=ret, 
		auc=acc))
}

```


















1.       Entire dataset > PNP with latent approach (no traits) > include citation counts > Return a csv of mammal species and probability of interaction with virus_genus = ‘Betacoronavirus’ as outcome variable > DallasMammalsCitations.csv
2.       Entire dataset > PNP with latent approach (no traits) > don’t include citation counts > Return a csv of mammal species and probability of interaction with virus_genus = ‘Betacoronavirus’ as outcome variable > DallasMammalsUncorrected.csv







```{r}

test <- formatPPobject(intMatFull)
out <- getReal(test, cv=FALSE)

topLinks <- cbind(test, suitability=out$predictions[,'suitability'])
topLinks <- topLinks[which(topLinks$presence == 0),]
topLinks <- topLinks[order(topLinks$suitability),]

retNoPredictors <- topLinks[which(topLinks$pathogen == 'Betacoronavirus'), c(1,2,7)]

write.csv(retNoPredictors, file='DallasDumbPredictions.csv')

```



















































# Betacoronavirus 

linksDF2 <- left_join(linksDF, virus[,c('vVirusNameCorrected', 'vFamily')],
	by=c('virus'='vVirusNameCorrected'))

linksCov <- filter(linksDF2, vFamily=='Coronaviridae')
linksCov2 <- filter(linksCov, presence==0)
linksCov3 <- linksCov2 %>% 
	group_by(host) %>%
	summarise(meanSuit=mean(suitability))

linksCov3 <-linksCov3[order(linksCov3$meanSuit),]


```











Small model

```{r}
# code below not edited. Models need to be trained first.
linksDFC <- data.frame(host=rownames(hvDati[[1]])[hvOut1[[1]]],
	virus = colnames(hvDati[[1]])[hvOut1[[2]]], 
	presence = hvOut1[[3]]$presence,
	suitability = hvOut1[[3]]$suitability)

linksDF <- linksDF[order(linksDF$suitability),]

linksDF2 <- left_join(linksDF, virus[,c('vVirusNameCorrected', 'vFamily')],
	by=c('virus'='vVirusNameCorrected'))

linksCov <- filter(linksDF2, vFamily=='Coronaviridae')
linksCov2 <- filter(linksCov, presence==0)
linksCov3 <- linksCov2 %>% 
	group_by(host) %>%
	summarise(meanSuit=mean(suitability))

sort(table(linksCov2$host))

```







































```{r}

#plug and play cross validated
auc <- vector()
for(i in 1:100){
  set.seed(i)
  auc[i] <- getReal(hvDati[[4]], cv=TRUE, folds=5)$auc
}

```


## Variable importance

```{r eval=FALSE, echo=FALSE}


getImportance <- function(column, pointsobject, n=100){
	require(ROCR)
  #all the data divided into presence (p) and background (bg)
  p <- pointsobject[(pointsobject[,'presence'] == 1), 4:ncol(pointsobject)]
  bg <- pointsobject[,4:ncol(pointsobject)]

	shufTrain <- function(bg, ind=1){
    accChange <- vector()
#   bg <- pointsobject[,4:ncol(pointsobject)]
    #shuffle interactions of variable i
    for(z in seq_len(n)){
      bg[,ind] <- bg[sample(1:nrow(vars),nrow(vars)),ind]
      fullmodel <- pp.gauss(p=bg[(pointsobject$presence == 1), ], 
				bgrd=bg, vbose=FALSE)
      redpreds <- predict.pp.gauss(fullmodel, x=bg)
      ret <- data.frame(presence=pointsobject[,'presence'], 
				suitability=redpreds)
      #auc calculation 
      pred <- prediction(predictions=ret[,2], labels=ret[,1])
      accChange[z] <- as.numeric(performance(pred, 'auc')@y.values)
		}
		return(c(mn=mean(accChange,na.rm=TRUE), sd=sd(accChange, na.rm=TRUE)))
	}

	varImp <- vector()
	tmp <- try(shufTrain(bg, ind=column))
	if(inherits(tmp, 'try-error')){
		varImp <- c(NA,NA)
	}else{
		varImp <- unlist(tmp)
  }
  #positive values indicate that variable permutation improved model performance
  #negative values indicate that variable permutation decreased model performance
  return(varImp)
}

```






### Obtain overall variable importance values

```{r}

library(doParallel)  
cl <- makeCluster(12, type="SOCK") 
registerDoParallel(cl)  
varImportance <- parLapply(cl, 1:175, getImportance, hvDati[[4]], n=100)
stopCluster(cl)



varImport <- do.call(rbind, varImportance)
varImport <- as.data.frame(varImport)
varImport$variable <- colnames(hvDati[[4]])[4:ncol(hvDati[[4]])]

```







```{r}

plotVarImport <- function(vdat, rows=nrow(vdat), 
	legX=0.01, legY=60,
	rowNames=NULL, colz=NULL, reduced=FALSE, scale=TRUE){
	vdat <- vdat[order(vdat$mn, decreasing=TRUE),]
	if(scale){
		vdat$mn <- vdat$mn - hvOut1$auc
		vdat$mn <- abs(vdat$mn - max(vdat$mn))
	}
	if(is.null(colz)){
		colorz <- c('purple', '#0099ff', '#14cad4', '#005B99', '#99005b', '#990000', '#ff0000', '#FF9999', '#FFCC99')
		colz <- colorz[as.numeric(as.factor(vdat$broadColorz))]
	}	
	if(reduced==FALSE){
		par(mar=c(4,0.5,0.5,0.5))
		plot(x=vdat$mn, y=1:nrow(vdat), pch=16, col=colz, las=1, yaxt='n', ylab='',
			xlim=c(min(vdat$mn-vdat$sd, na.rm=TRUE),max(vdat$mn+vdat$sd, na.rm=TRUE)),
			xlab='change in AUC')
		segments(x0=vdat$mn-vdat$sd, x1=vdat$mn+vdat$sd, y0=1:nrow(vdat), col=colz)
		abline(v=hvOut1$auc, col=grey(0.5,0.9), lwd=2)
		points(x=vdat$mn, y=1:nrow(vdat), col=colz, pch=16, cex=1.1)

		points(x=vdat$mn[which(vdat$mn-vdat$sd > hvOut1$auc)], 
			y=which(vdat$mn-vdat$sd > hvOut1$auc), 
			col='red', bg=colz[which(vdat$mn-vdat$sd > hvOut1$auc)], pch=21)
		leg <- data.frame(cat=unique(vdat$broadColorz), 
			ind=unique(as.numeric(as.factor(vdat$broadColorz))), 
			col=substr(unique(colz), start=1, stop=7))
		leg$col <- as.character(leg$col)
		leg$cat <- as.character(leg$cat)
		leg <- leg[order(leg$ind),]
		legend(legX, legY, col=leg$col, leg$cat, pch=16, bty='n', pt.cex=1.5)
	}
	if(reduced){
		vdat2 <- vdat %>% group_by(broadColorz) %>% summarise(mnn=mean(mn), sdd=mean(sd))
		vdat2 <- vdat2[order(vdat2$mnn),]
		par(mar=c(4,9,0.5,0.5))
		plot(x=vdat2$mnn, y=1:nrow(vdat2), pch=16, col=colorz, las=1, yaxt='n', ylab='',
			xlim=c(min(vdat2$mnn-vdat2$sdd, na.rm=TRUE),max(vdat2$mnn+vdat2$sdd, na.rm=TRUE)),
			xlab='change in AUC')
		axis(2, at=1:nrow(vdat2), labels=unique(vdat2$broadColorz), las=1)
		segments(x0=vdat2$mnn-vdat2$sdd, x1=vdat2$mnn+vdat2$sdd, y0=1:nrow(vdat2), col=colorz)
		abline(v=hvOut1$auc, col=grey(0.5,0.9), lwd=2)
		points(x=vdat2$mnn, y=1:nrow(vdat2), col=colorz, pch=16, cex=1.1)
	}
}

```











### Identifying likely missing links

```{r}

topLinks <- linksDF[which(linksDF$suitability > 1e100 & linksDF$presence == 0),]

missedLinks <- linksDF[which(linksDF$suitability == 0 & linksDF$presence == 1),]

```




